Geometrization of characters of tori over non-Archimedean local fields

### 2013-07-27 shanghai.tex
Working on base change for Gr(Neron)

\newcommand{\FF}{{\mathbb{F}}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\NN}{{\mathbb{N}}}
\newcommand{\CC}{{\mathbb{C}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\RR}{{\mathbb{R}}}
\newcommand{\EE}{\mathbb{\bar Q}_\ell}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\pK}{\mathfrak{p}_K}
\newcommand{\Zp}{\mathbb{Z}_p}
\newcommand{\Qp}{\mathbb{Q}_p}
%\newcommand{\Fp}{\mathbb{F}_p}
\newcommand{\bFq}{\bar{k}}
\newcommand{\Fq}{k}
\newcommand{\Fqm}{k_m}
%\newcommand{\bFp}{{\mathbb{\bar F}_p}}

@@

\newcommand{\cdef}[1]{{{\color{cyan}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\GC}{{\mathcal{G\hskip-0.8pt C}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CS_0(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}

@@

Let $\GN{T}_{d}$ be the group scheme over $\Fq$ produced by applying
the Greenberg functor to $\TT_{d}$. (See \cite{Greenberg:2} and
\cite[Ch. 9, \S 6]{BLR} for background on the Greenberg functor.)
%Then $\GN{T}_d(\Fq) = \TT(\OK/\pK^d)$.

@@

We may now assemble the group $\Fq$-schemes $\TT_{d}$ into one
glorious group pro-scheme over $\Fq$:
\begin{equation}\label{projlim}
  \GN{T} \ceq \varprojlim_{d\in \NN} \GN{T}_{d}.
\end{equation}
This is the \cdef{Greenberg transform of the lft-N\'eron model of $T$}.

@@

\begin{proposition}
Let $K'/K$ be a finite Galois extension and let $T'$ be an algebraic torus over $K'$... 
\Clifton{This next part is very much under construction.}
The functor $\CS(\GN{T}) \to \CS(\GN{T}')$ defined above induces a group homomorphism 
that commutes, through trace of Frobenius, with ...
\[
\begin{tikzcd}
\CS(\GN{T}) \dar & \CSiso{\GN{T}} \rar{t_T} \dar & \Hom_\text{ad}(T(K),\EEx) \dar \\
\CS(\GN{T}') & \CSiso{\GN{T}'} \rar{t_{T'}} & \Hom_\text{ad}(T'(K),\EEx)
\end{tikzcd}
\]
\end{proposition}

\begin{proof}
\end{proof}

@@

>>>>

\newcommand{\FF}{{\mathbb{F}}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\NN}{{\mathbb{N}}}
\newcommand{\CC}{{\mathbb{C}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\RR}{{\mathbb{R}}}
\newcommand{\EE}{\mathbb{\bar Q}_\ell}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\pK}{\mathfrak{p}_K}
\newcommand{\OL}{\mathcal{O}_L}
\newcommand{\Zp}{\mathbb{Z}_p}
\newcommand{\Qp}{\mathbb{Q}_p}
%\newcommand{\Fp}{\mathbb{F}_p}
\newcommand{\bFq}{\bar{k}}
\newcommand{\Fq}{k}
\newcommand{\Fqm}{k_m}
%\newcommand{\bFp}{{\mathbb{\bar F}_p}}

@@

\newcommand{\cdef}[1]{{{\color{cyan}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\TL}{\underline{T_L}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\GC}{{\mathcal{G\hskip-0.8pt C}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CS_0(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}

@@

Let $\GN{T}_{d} = \Gr(\TT_d)$ be the group scheme over $\Fq$ produced by applying
the Greenberg functor to $\TT_{d}$. (See \cite{Greenberg:2} and
\cite[Ch. 9, \S 6]{BLR} for background on the Greenberg functor.)
%Then $\GN{T}_d(\Fq) = \TT(\OK/\pK^d)$.

@@

We may now assemble the group $\Fq$-schemes $\TT_{d}$ into one
glorious group pro-scheme over $\Fq$:
\begin{equation}\label{projlim}
  \GN{T} \ceq \varprojlim_{d\in \NN} \GN{T}_{d}.
\end{equation}
This is the \cdef{Greenberg transform of the lft-N\'eron model of $T$}; we will also write $\Gr$ for this functor on $\OK$-schemes.

@@

Let $L/K$ be a finite Galois extension and let $K'/K$ be the maximal unramified subextension, with residue field $k'$.  If $T$ is a torus over $K$, write $\TL$ for the N\'eron model of the base change $T_L$.  Then there is a canonical isomorphism
\[
\Res_{k'/k} \Gr(\TL) \cong \Gr(\underline{\Res_{L/K}(T_L)}).
\]
The natural inclusion $T \hookrightarrow \Res_{L/K} T_L$ induces a map of $k$-schemes
\[
\GN{T} \hookrightarrow \Res_{k'/k} \Gr(\TL)
\]
and thus a map on character sheaves.

\begin{proposition}
The following diagram commutes:
\[
\begin{tikzcd}
\CSiso{\Res_{k'/k} \Gr(\TL)} \arrow[two heads]{r} \dar & \CSiso{\GN{T}} \dar\\
\Hom(T(L), \EEx) \arrow[two heads]{r} & \Hom(T(K), \EEx)
\end{tikzcd}
\]
\end{proposition}

We now translate the norm map $T(L) \to T(K)$ to character sheaves.  There is an open immersion
\[
\TT \times_{\OK} \OL \to \TL
\]
induced by the canonical isomorphism of generic fibers \cite[p. 1]{Chai}.  After taking Greenberg transforms,
we obtain a diagram
\[
\begin{tikzcd}
\Gr(\TT) \dar & \Gr(\TT) \times_{\Gr(\OK)} \Gr(\OL) \lar{\alpha} \rar{\beta} \dar & \Gr(\TL) \dar \\
\Spec{k} & \Spec{k'} \lar \rar & \Spec{k'}
\end{tikzcd}
\]
In the case that $L/K$ is unramified, the left hand square is a pullback square and the right hand horizontal maps are isomorphisms.  In general, we may define a functor
\begin{align*}
\CS(\GN{T}) &\to \CS(\Gr(\TL)) \\
\cs{L} &\mapsto \beta_* \alpha^* \cs{L}.
\end{align*}

\begin{proposition}
The following diagram commutes:
\[
\begin{tikzcd}
\CSiso{\GN{T}} \rar{\beta_* \circ \alpha^*} \dar & \CSiso{\Gr(\TL)} \dar \\
\Hom(T(K), \EEx) \rar{\Nm} & \Hom(T(L), \EEx)
\end{tikzcd}
\]
\end{proposition}

@@

\bib{Chai}{article}{
   author={Chai, C.L.},
   title={N\'eron models for semiabelian varieties: congruence and change of base field.},
   journal={Asian J. Math.},
   volume={4},
   number={4},
   date={2000},
   pages={715--736}}
